---
title: "Homework 04 Submission"
author: "Jimmy Liu"
date: "2019-10-08"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

\newpage

```{r, echo = F, warning = F, message = F}
library(gapminder)
library(tidyverse)
library(DT)
```

# Exercise 1

## Widen gapminder dataset by creating a lifeExp column for each country
```{r, tidy=TRUE, tidy.opts=list(width.cutoff=110)}
gapminder_wide <- gapminder %>% 
  filter(country %in% c("Canada", "China")) %>% # Filter for rows with Canada or China
  pivot_wider(id_cols = c(country, year), # Keep countries and year columns
              values_from = lifeExp, # Cells of new columns will contain lifeExp values
              names_from = country) # Newly generated columns are named by the values
                                    # of the country column

gapminder_wide %>% 
  knitr::kable() # Present the transformed tibble in a nice table
```

## Create a scatterplot of life expectancies in Canada vs China
```{r, fig.height=3, fig.width=5}
gapminder_wide %>% 
  ggplot(aes(x = Canada, y = China)) + # Specify the x and y variables for ggplot
  geom_point() + # Add point aesthetics for each data point
  xlab("Canada life expectancy") + # label x axis
  ylab("China life expectancy") # label y axis
```

## Re-lengthen the widen gapminder dataset
```{r}
gapminder_wide %>% 
  pivot_longer(cols = (-year), # Select all columns except year to be collapsed
               names_to = "country", # Collapse Canada and China columns to 
                                     # a column called "country"
               values_to = "lifeExp") %>% # Values of Canda and China columns
                                          # will be under a column called "lifeExp" 
  DT::datatable() # Present the transformed tibble in a nice table
```

# Exercise 2

## Widen gapminder dataset by creating lifeExp and gdpPerCap columns for each country
```{r}
gapminder_wide <- gapminder %>% 
 filter(country %in% c("Canada", "China")) %>% # Filter for rows with Canada or China
 pivot_wider(id_cols = c(country, year), # Keep countries and year columns
             names_from = country, # Newly generated columns are named by
                                   # the values of the country column
             names_sep = "_", # Each new column name is separated by "_"
             values_from = c(lifeExp, gdpPercap)) # Cells of the new columns will contain
                                                  # either lifeExp or gdpPerCap values

gapminder_wide %>% 
  knitr::kable() # Present the transformed tibble in a nice table
```

### Re-lengthen the widen gapminder dataset
```{r}
gapminder_wide %>% 
  pivot_longer(cols = (-year), # Collapse all columns except the year column
               names_sep = "_", # Split the column names by "_"
               names_to = c(".value", "country")) %>% # Collapse selected columns
                                                      # and create 3 new columns
  DT::datatable() # Present the transformed tibble in a nice table
```


# Exercise 3

```{r, warning = F, message = F, echo = F}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```

```{r, warning = F, message = F, eval = F}
guest <- read_csv("../data/wedding/attend.csv")
email <- read_csv("../data/wedding/emails.csv")
```

## Merge the email column to the guest dataset
```{r}
guest %>% 
  rename("guest" = name) %>% # rename the name variable to "guest"
  left_join((email %>% 
               separate_rows("guest", sep = ", ")), # split values in the guest 
                                                    # variable into separate rows
            by = "guest") %>% # merge by the guest variable
  select(party, guest, email) %>% # Only keep the party, guest and email columns
  DT::datatable() # Present the new tibble in a nice table
```

## Report individuals with known emails but not on guest list
```{r}
email %>% 
  rename("name" = guest) %>% # rename the guest variable to "name"
  select(name) %>% # keep the name variable only
  separate_rows("name", sep = ", ") %>% # split values in the name 
                                        # variable into separate rows
  setdiff(guest %>% # find values in the email tibble not 
                    # in the guest tibble
            select(name)) %>%
  knitr::kable() # present the new tibble in a nice table
```

## Report everyone in the guest and email datasets
```{r}
email %>% 
  rename("name" = guest) %>% # rename the guest variable to "name"
  select(name) %>% # keep the name variable only
  separate_rows("name", sep = ", ") %>% # split values in the name variable into separate rows
  union(guest %>% # join all values found in the email and guest tibble into a single tibble
            select(name)) %>% 
  DT::datatable() # present the new tibble in a nice table
```

