---
title: "HW02_dplyr"
output:
  github_document:
    toc: true
    toc_depth: 3
---

```{r, warning=F, echo=F, message=F}
library(DT)
library(gapminder)
library(tidyverse)
```
## Exercise 1

### Display gapminder dataset for 3 countries in 1970s

```{r, warning=F}
gapminder %>% 
  filter(country %in% c("Austria", "Belgium", "Australia"),
         year %in% c(1970:1979)) %>% 
  as.tibble() %>% 
  knitr::kable()
```

### Display the above table with only country and GDP Per Capita information

```{r, warning=F}
gapminder %>% 
  filter(country %in% c("Austria", "Belgium", "Australia"),
         year %in% c(1970:1979)) %>%
  select(country, gdpPercap) %>% 
  as.tibble() %>% 
  knitr::kable()
```

## Display entries in which countries experienced decrease in life expectancies

```{r, warning=F}
gapminder %>%
  group_by(country) %>% 
  mutate(lifeExpChanges = c(0,diff(lifeExp))) %>% 
  filter(lifeExpChanges < 0) %>% 
  as.tibble() %>% 
  DT::datatable()
```

### Display entries in which countries had their highest GDP per capita

```{r, warning=F}
gapminder %>% 
  group_by(country) %>% 
  filter(gdpPercap == max(gdpPercap)) %>% 
  as.tibble() %>% 
  DT::datatable()
```

### Plot life expectancies vs GDP per capita in Canada

```{r, warning=F}
gapminder %>% 
  filter(country == "Canada") %>% 
  ggplot(aes(x = log10(gdpPercap), y = lifeExp)) + 
  geom_point()
```

## Exercise 2

### List all the continents surveyed in the gapminder dataset

```{r}
as.character(unique(gapminder$continent))
```

### Display the number of countries surveyed per continent

```{r}
as.data.frame(table(gapminder$continent)) %>% 
  mutate(Countries = Freq/12) %>% 
  ggplot(aes(x = Var1, y = Countries)) + 
  geom_bar(stat = "identity") +
  xlab("Continent")
```

### Show the minimum and maximum population value in the gapminder dataset

```{r}
paste("Minimum population value: ", min(gapminder$pop))
paste("Maximum population value: ", max(gapminder$pop))
```

### Summary statistics of the population continuous variable in the gapminder dataset

```{r}
summary(gapminder$pop)
```

## Exercise 3

### Plotting the correlation of Petal width and length across three different flower species from the iris dataset

```{r}
iris %>% 
  ggplot(aes(x = Petal.Length, y = Petal.Width, color = Species)) +
  geom_point()
```

### Plotting the horsepower of different car models from the mtcars dataset

```{r}
mtcars %>% 
  ggplot(aes(x = rownames(mtcars), y = hp)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  xlab("Car Models") +
  ylab("Horsepower")
```


