---
title: "hw05"
author: "Jimmy Liu"
date: "16/10/2019"
output: 
  pdf_document:
   toc: true
   toc_depth: 3
   number_sections: true
---

\newpage

# Purpose of here() package

The here() package is a convenient tool that enables easy sharing of R scripts. In the case when a project session is not opened, any R scripts assumes a default working directory which can be slightly different depending on the users accessing the scripts. Moreover, it is impractical to begin a script with setwd() to specify the path to the current working directory, as the directory structure varies between different devices and operating systems. These issues render R scripts unable to run smoothly if it requires access to files in particular directories. The here() package addresses the issues by looking for .Rproj project files in the current or parent directories of the R script being accessed and assumes the directory where .Rproj files are found to be the current working directory.

```{r, message = F, echo = F}
library(gapminder)
library(tidyverse)
library(here)
library(scales)
```

# Explore factors with gapminder dataset

The gapminder dataset is filtered to remove observations in Oceania and summarized to find the highest GDP per Capita values in each continent. The level corresponding to "Oceania" in the continent factor is removed.
```{r}
p <- gapminder %>% 
  filter(continent != "Oceania") %>% 
  droplevels() %>% 
  group_by(continent) %>% 
  summarize(max_gdpPercap = max(gdpPercap))
```

A barplot is created using the above tibble while maintaining the original factor levels in the continent factor.
```{r, fig.width = 5, fig.height = 3}
p %>% 
  ggplot(aes(x = continent, y = max_gdpPercap)) +
  geom_bar(stat = "identity") +
  ylab("Maximum GDP per Capita") +
  xlab("Continent") +
  scale_y_continuous(labels = scales::dollar_format())
```

A barplot is created with the same tibble, but now the continent factor is rearranged based on their highest GDP per Capita values.
```{r, fig.width = 5, fig.height = 3}
p %>% 
  ggplot(aes(x = fct_reorder(continent, max_gdpPercap), y = max_gdpPercap)) +
  geom_bar(stat = "identity") +
  ylab("Maximum GDP per Capita") +
  xlab("Continent") +
  scale_y_continuous(labels = scales::dollar_format())
```

# Exploring file input and output

Modify the gapminder dataset by summarizing the average life expectancies across all countries and years in each continent and subsequently rearranging the continent factor by their average life expectancies.
```{r}
# Modify the gapminder dataset
p <- gapminder %>% 
  group_by(continent) %>% 
  summarize(lifeExp_mean = mean(lifeExp))
p$continent <- fct_reorder(p$continent, p$lifeExp_mean)

p
```

Write out the modified gapminder as a csv. The csv file is then re-imported back and the dataset is printed to demonstrate that read_csv() ran successfully.
```{r, message = F}
# Write out csv
write_csv(p, here("hw05", "gapminder_modified.csv"))

# Read back csv
read_csv(here("hw05", "gapminder_modified.csv"))
```

# Improving visualization design

From the visualization design lecture,  we learned it is best to keep figures and plots as simple as possible to prevent particular features from distracting the audience. From the original plot, the grid lines and gray background was removed. The self-explainatory axis and legend titles were removed. A single plot title was included to clarify the variables used in the plot. Overall the revised plot is much cleaner and allows the audience to focus on the trends and data shown rather than distracted by unnecessary features in the original plot.
```{r, fig.width = 5, fig.height = 3}
# Original Plot
gapminder %>%
  group_by(continent, year) %>% 
  summarize(mean_lifeExp = mean(lifeExp)) %>% 
  ggplot(aes(x = year, y = mean_lifeExp, color = continent)) +
  geom_line() +
  geom_point()
```

```{r, fig.width =5 , fig.height = 3, message = F}
# Revised Plot
gapminder %>%
  group_by(continent, year) %>% 
  summarize(mean_lifeExp = mean(lifeExp)) %>% 
  ggplot(aes(x = year, y = mean_lifeExp, color = continent)) +
  geom_smooth(se = F) +
  theme(panel.grid.major = element_blank(), # removing grid lines
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"),
        legend.title = element_blank(), 
        axis.title.x=element_blank(), 
        axis.title.y=element_blank()) +
  ggtitle("Global Life Expectancy Trends from 1952")
```

# Writing figures to file

A bubble plot is created from the gapminder dataset to study the correlation between GDP per Capita and life expectancy in Africa. The size of the datapoints represent the population of each country in 1972. The corresponding plot is saved as "gapminder_plot_ggsave.jpg" using the ggsave() function.
```{r, message = F, fig.width = 5, fig.height =3}
p <- gapminder %>% 
  filter(continent == "Africa",
         year == "1972") %>% 
  ggplot(aes(gdpPercap, lifeExp, size = pop, fill = continent)) +
  geom_point(alpha = 0.5, shape = 21) +
  scale_x_log10(labels = scales::comma_format()) +
  scale_size_area() +
  guides(fill = FALSE) +
  xlab("GDP Per Capita") +
  ylab("Life Expectancy (years)") +
  scale_x_continuous(labels = scales::dollar_format())
ggsave(here("hw05", "gapminder_plot_ggsave.jpg"), p)
plot(p)
```

